\mytitile\\~

### 2019-10-09 document.tex
New appendics

\usepackage{graphicx}% Include figure files
%\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{fixme}
%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{mathptmx}
\usepackage{lipsum}
\usepackage{amsmath}
\usepackage{physics}
\usepackage{xparse}
\usepackage{bbm}
\usepackage{hyperref}
%\usepackage{multirow}
%\usepackage{makecell}
\graphicspath{{Pictures/}}

\renewcommand*{\figureautorefname}{Fig.}

\newcommand{\mytitile}{Effects of strong interaction between light and matter in a superconducting artificial molecule}
\begin{document}
	\preprint{AIP/123-QED}

@@

	\maketitle
\section{Introduction}
Over the past twenty years, superconducting artificial atoms (SAA) were used in numerous experiments confidently demonstrating the validity of fundamental quantum mechanical laws\cite{you2011atomic}. Due to synthetic nature of SAA, their Hamiltonians can be pre-designed and engineered, and this makes them a particularly versatile tool for studies in quantum optics. Notably, their parameters may be varied in such a wide range that novel physical effects previously inaccessible for natural systems may be directly observed. For example, the strong coupling regime when the coupling strength between light and a quantum system approaches 1\% of their frequency has been a major milestone for atoms, molecules, etc and was only reached using high-finesse cavities. In contrast, superconducting quantum systems have rapidly attained this landmark using \textit{circuit QED}\cite{wallraff2004strong} just in a few years after their first discovery\cite{chiorescu2004coherent} and have already surpassed other strongly-coupled systems in terms of coherence\cite{forn2019ultrastrong}. 
Moreover, SAA do not even require confined radiation to implement strong coupling with light: they may be coupled unprecedently strong with the free-propagating EM-waves in the on-chip 1D waveguides\cite{astafiev2010resonance} without having to use cavities at all. The Rabi frequency in this case may reach 50\% of the transition frequency\cite{deng2015observation} which is already the ultra-strong coupling regime.

In general, a quantum system exposed to strong external driving can experience various quantum phenomena explained by resonant single- and multi-photon transitions between energy levels. The most known examples are coherent population trapping, electromagnetically induced transparency (EIT)\cite{boller1991observation}, stimulated Raman adiabatic passage\cite{bergmann1998coherent} and Autler-Townes (A-T) splitting\cite{autler1955stark}.

In this work, we are studying effects of light-matter interaction that occur in a system of two coupled artificial atoms: a superconducting artificial molecule (SAM). In our experiment, we use two transmons\cite{koch2007charge} interacting with each other through a cavity bus\cite{majer2007coupling}. Microwave radiation is applied to this system through a coplanar on-chip antenna and transmon states may be monitored by dispersive readout using the same coupling cavity \cite{chow2010detecting}.

It turns out that due to the strong interaction of SAM with the microwave drive, multiphoton transitions and the Autler-Townes effect are playing the crucial role in the phenomena that we have discovered. The A-T splitting has been observed before in atomic\cite{picque1976direct}, molecular systems\cite{tamarat1995pump}, quantum dots\cite{xu2007coherent}, and for SAAs as well\cite{baur2009measurement, sillanpaa2009autler, novikov2013autler, suri2013observation, peng2018vacuum}. However, the previous studies have  only involved just a single SAA and have demonstrated only the standard spectral signatures known from the quantum optics; instead, a diatomic SAM allows to observe some qualitatively new spectral manifestations of the effect.

Remarkably, besides its fundamental value, this experiment may be useful for future quantum computers: a quantum engineer should consider the observed behaviours as unwanted, and thus careful control of the driving power must be employed to avoid them (for instance, as we will show, the bSWAP gate\cite{poletto2012entanglement} is directly affected).

In  cause of the AT effect is the dressing of the atomic levels by strong EM radiation. This dressing The new energy state called dressed states. If an additional weak probing field is applied, the spectra of the strongly driven system can be probed. In particular the dressed state model nicely explains the emergence of the Mollow triplet\cite{newstein1968spontaneous} spectrum, which consists of a central peak with two side bands separated by the Rabi frequency.

\begin{figure}
	\includegraphics[width=\linewidth]{intro_scheme}
	\caption{Illustrating the AT splitting in the rotating frame ($\hbar=1$ here). \textbf{(a)} The three-level system is driven strongly by a ``coupler'' tone of amplitude $\Omega$ at frequency $\omega_{01}-\Delta$. It induces Rabi oscillations between $\ket{0}$ and $\ket{1}$ at the generalized Rabi frequency $\Omega_R = \sqrt{\Omega^2 + \Delta^2}$. \textbf{(b)} Moving to the frame rotating with the drive, we effectively change the $\ket{0}\rightarrow \ket{1}$ transition frequency to $\Delta$. However, when RWA is applied and Hamiltonian is re-diagonalized, the splitting between two lowest levels (dressed states $\ket{a}$ and $\ket{b}$) becomes $\Omega^R$. Now, the experimenter may observe a doublet transition from these levels to the state $\ket{2}$ with frequencies $\omega_{12}+(\Delta \pm \Omega^R)/2$.} 
\end{figure}

@@

A sufficiently large anharmonicity is needed to prevent qubit operations from exciting other transitions in the system. Therefore, we will use two- and three-level approximations for our calculations

In this article will be shown how the spectrum of the system depends on drivers on different qubits and explicitly shown the dressed states.%\autoref{sec:level1}

\section{Methods}
\subsection{Theoretical description of the system}

We investigate a system of two interacting transmons. The transmon represents two Josephson junctions (dc-SQUID) which are shunted by an additional large capacitance(compared to other capacitances in the circuit). This dc-SQUID setup allows for the tuning of the Josephson energy $E_J = E_J^{max}|\cos(\pi \frac{\Phi}{\Phi_0})|$  by
means of an external magnetic flux $\Phi$. The transmon can be considered as an oscillator with a quartic perturbation describing the leading-order anharmonicity. The system of two transmons and their connection forms a level system.

In this section we analyse theoretically two qubit system.
 In our study we used transmons.The transmon Hamiltonian is $H = 4\cdot E_C(n-n_g)^2 -E_J\cos(\phi)$, where $E_c $ and $E_J$ -charge and Josephson energy respectively. The symbols $n$ and $\phi$ denote the number of Cooper pairs transferred between the islands and the
 gauge-invariant phase difference between the superconductors,
 respectively.
 When the ratio $\frac{E_C}{E_J}$ is large the perturbation approach can
 be employed. After the expansion of the cosine for small angles and neglecting of the small terms the Hamiltonian can be
 viewed as a harmonic oscillator with a quartic perturbation describing the leading-order anharmonicity.
 After conversion to dimensionless variables and neglecting small terms and set $\hbar = 1$, we move to the new Hamiltonian of transmon. 
\begin{equation}
H_{tr}= \omega_{tr} b^{\dagger}b +\frac{1}{2}\alpha b^{\dagger}b(b^{\dagger}b-1)
\end{equation}
Where $\alpha=-E_C$, $\omega_{tr}=\sqrt{8E_CE_J}+\alpha$.

The implementations of the transitions between the eigenstates is achieved by external microwave driving through a capacitively coupled transition line. In the notations used above, the pump Hamiltonian can be written generically as $ 	H_{dr} = -2e n\beta V_{dr}(t)=\Omega(t)(b+b^{\dagger})$.
Where $2en$ -- the charge on island, $\beta$ -- dimensionless coefficient depending on the geometry of the system and the ratio of capacities, $V_{dr}(t)$ -- the voltage on the external source.

@@

From the simulation follows that a two-level approximation for first qubit and three-level approximation for second is sufficient to observe the effect we are interested in. The Hamiltonian in matrix form is given by the following expression:
\begin{widetext}
	\begin{equation}\label{matrix} 
	\text{$H^R$} = \left(\begin{matrix}
	0 & \text{$\Omega_1/2 $} & \text{$\Omega _2$} e^{i \text{$\delta_{\omega}$} t}/2 & 0 & 0 & 0 \\
	\text{$\Omega_1/2 $} & \text{$-\Delta_1 $} & J & \text{$\Omega_2$} e^{i \text{$\delta_{\omega}$} t}/2 & 0 & 0 \\
	\text{$\Omega_2$} e^{-i \text{$\delta_{\omega} $} t}/2 & J & -\text{$\Delta_2 $} &
	\text{$\Omega_1$/2} & \sqrt{2} \text{$\Omega_2$} e^{i \text{$\delta_{\omega}$} t}/2 & 0 \\
	0 & \text{$\Omega_2$} e^{-i \text{$\delta_{\omega} $} t}/2 & \text{$\Omega_1/2$} &
	\text{$-\Delta_2 -\Delta_1$} & \sqrt{2} J & \sqrt{2} \text{$\Omega_2$} e^{i \text{$\delta_{\omega}
			$} t}/2 \\
	0 & 0 & \sqrt{2} \text{$\Omega_2$} e^{-i \text{$\delta_{\omega}$} t}/2 & \sqrt{2} J &
	\text{$\alpha$}-2 \text{$\Delta_2$} & \text{$\Omega_1/2$} \\
	0 & 0 & 0 & \sqrt{2} \text{$\Omega_2$} e^{-i \text{$\delta_{\omega} $} t}/2 & \text{$\Omega_1/2
		$} & \text{$\alpha$}-2 \text{$\Delta_2$}- \text{$\Delta_1$} \\
	\end{matrix}\right)
	\end{equation}
\end{widetext}
Where $\Omega_1$, $\Omega_2$ - amplitudes of the drives, $\Delta_1 = \omega_d-\omega_{tr1}=0$,$\Delta_2 = \omega_d-\omega_{tr2}$.
Let's consider the Hamiltonian (\autoref{matrix}) in details. Divide it into three parts: time-dependent $(V_t)$, the part responsible for the interaction between the qubits $(V_J)$ and all the rest $(H_0)$.
\begin{equation}
	H^R=H_0+V_J+V_t
\end{equation}
The $\Omega_1\gg \Omega_2,\Delta_2\gg J$, so the driven field on the first qubit is strong and form the dressed states defined by the Hamiltonian $H_0$.
The Hamiltonian $H_0$ is station and we can find eigenvectors and eigenvalues. The drive on the second qubit is weak with very low intensity, so we can neglect any perturbation introduced by second drive on the position, width os the dressed states.
Firstly, we consider the system of non-interacting qubits $(J=0)$. In this case some transitions are prohibited and the splitting we are interested in is not observed. 
Therefore a non-zero interaction between qubits is necessary to observe AT effect.

After that according to the perturbation theory (with perturbation $V_J$) to $H_0$ we found the first-order corrections to the wave functions. We can use Fermi Golden Rule to find the probability of transition.
\appendix

\subsection{The standard Autler-Townes effect}

\section{Theoretical description of two-photon process}

>>>>

\usepackage{graphicx}% Include figure files
%\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines
\usepackage{hyperref}
\usepackage{kbordermatrix}% http://www.hss.caltech.edu/~kcb/TeX/kbordermatrix.sty
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{mathptmx}
\usepackage{lipsum}
\usepackage{amsmath}
\usepackage{physics}
\usepackage{xparse}
\usepackage{bbm}
\usepackage{nicematrix}
%\usepackage{multirow}
%\usepackage{makecell}
\graphicspath{{Pictures/}}

\newcommand{\mytitile}{Effects of intense light-matter interaction in a superconducting artificial molecule}
\begin{document}
	\preprint{AIP/123-QED}

@@

	\maketitle
\section{Introduction}
Over the past twenty years, significant progress has been achieved in the research of the superconducting quantum circuits.

We investigate a system of two interacting transmons. The transmon\cite{koch2007charge} represents two Josephson junctions (dc-SQUID) which are shunted by an additional large capacitance(compared to other capacitances in the circuit). This dc-SQUID setup allows for the tuning of the Josephson energy $E_J = E_J^{max}|\cos(\pi \frac{\Phi}{\Phi_0})|$  by
means of an external magnetic flux $\Phi$. The transmon can be considered as an oscillator with a quartic perturbation describing the leading-order anharmonicity. The system of two transmons and their connection forms a level system.
A sufficiently large anharmonicity is needed to prevent qubit operations from exciting other transitions in the system. Therefore, we will use two- and three-level approximations for our calculations

A quantum system exposed to external driving, can experience a number of quantum phenomena due to the resonant transitions between energy levels. The most known examples are coherent population trapping, electromagnetically induced transparency (EIT)\cite{boller1991observation}, stimulated Raman adiabatic passage\cite{bergmann1998coherent} and Autler-Townes (AT) splitting\cite{autler1955stark} which can happen in systems with just three energy levels. 
The main attention in our work is paid to the last effect, which related to the dressed states.
The AT effect has been studied in atomic\cite{picque1976direct} and molecular systems\cite{tamarat1995pump}, quantum dots\cite{xu2007coherent}, and superconducting qubits\cite{li2011decoherence, novikov2013autler}.

If the driving field is strong the interaction between the system and the field become significant and leads to oscillates with Rabi frequency. The new energy state called dressed states. If an additional weak probing field is applied, the spectra of the strongly driven system can be probed. In particular the dressed state model nicely explains the emergence of the Mollow triplet\cite{newstein1968spontaneous} spectrum, which consists of a central peak with two side bands separated by the Rabi frequency.

@@

In this article will be shown how the spectrum of the system depends on drivers on different qubits and explicitly shown the dressed states.\autoref{sec:level1}

\section{Methods}
\subsection{Theoretical description of the system}
In this section we analyse theoretically two qubit system.
 In our study we used transmons.The transmon Hamiltonian is $H = 4\cdot E_C(n-n_g)^2 -E_J\cos(\phi)$, where $E_c $ and $E_J$ -charge and Josephson energy respectively. The symbols $n$ and $\phi$ denote the number of Cooper pairs transferred between the islands and the
 gauge-invariant phase difference between the superconductors,
 respectively.
 When the ratio $\frac{E_C}{E_J}$ is large the perturbation approach can
 be employed. After the expansion of the cosine for small angles and neglecting of the small terms the Hamiltonian can be
 viewed as a harmonic oscillator with a quartic perturbation describing the leading-order anharmonicity.
 After conversion to dimensionless variables and neglecting small terms and set $\hbar = 1$, we move to the new Hamiltonian of transmon. 
\begin{equation}
H_{tr}= \omega_{tr} b^{\dagger}b +\frac{1}{2}\alpha b^{\dagger}b(b^{\dagger}b-1)
\end{equation}
Where $\alpha=-E_C$, $\omega_{tr}=\sqrt{8E_CE_J}+\alpha$.

The implementations of the transitions between the eigenstates is achieved by external microwave driving through a capacitively coupled transition line. In the notations used above, the pump Hamiltonian can be written generically as $ 	H_{dr} = -2e n\beta V_{dr}(t)=\Omega(t)(b+b^{\dagger})$.
Where $2en$ -- the charge on island, $\beta$ -- dimensionless coefficient depending on the geometry of the system and the ratio of capacities, $V_{dr}(t)$ -- the voltage on the external source; $n =\frac{1}{\sqrt{2}} (\frac{E_J}{8E_C})^{1/4}(b+b^{\dagger})$.

@@

From the simulation follows that a two-level approximation for first qubit and three-level approximation for second is sufficient to observe the effect we are interested in. The Hamiltonian in matrix form is given by the following expression:

\renewcommand{\kbldelim}{(}% Left delimiter
\renewcommand{\kbrdelim}{)}% Right delimiter
\begin{widetext}
	\begin{equation}\label{matrix} 
	\text{$H^R$} = \kbordermatrix{
	\ket{00} & \ket{10} & \ket{01} & \ket{11} & \ket{02} & \ket{12} \\
	0 & \text{$\Omega_1/2 $} & \text{$\Omega _2$} e^{i \text{$\delta_{\omega}$} t}/2 & 0 & 0 & 0 \\
	\text{$\Omega_1/2 $} & \text{$-\Delta_1 $} & J & \text{$\Omega_2$} e^{i \text{$\delta_{\omega}$} t}/2 & 0 & 0 \\
	\text{$\Omega_2$} e^{-i \text{$\delta_{\omega} $} t}/2 & J & -\text{$\Delta_2 $} &
	\text{$\Omega_1$/2} & \sqrt{2} \text{$\Omega_2$} e^{i \text{$\delta_{\omega}$} t}/2 & 0 \\
	0 & \text{$\Omega_2$} e^{-i \text{$\delta_{\omega} $} t}/2 & \text{$\Omega_1/2$} &
	\text{$-\Delta_2 -\Delta_1$} & \sqrt{2} J & \sqrt{2} \text{$\Omega_2$} e^{i \text{$\delta_{\omega}
			$} t}/2 \\
	0 & 0 & \sqrt{2} \text{$\Omega_2$} e^{-i \text{$\delta_{\omega}$} t}/2 & \sqrt{2} J &
	\text{$\alpha$}-2 \text{$\Delta_2$} & \text{$\Omega_1/2$} \\
	0 & 0 & 0 & \sqrt{2} \text{$\Omega_2$} e^{-i \text{$\delta_{\omega} $} t}/2 & \text{$\Omega_1/2
		$} & \text{$\alpha$}-2 \text{$\Delta_2$}- \text{$\Delta_1$} \\
}
	\end{equation}
\end{widetext}
Where $\Omega_1$, $\Omega_2$ - amplitudes of the drives, $\Delta_1 = \omega_d-\omega_{tr1}=0$,$\Delta_2 = \omega_d-\omega_{tr2}$.
Let's consider the Hamiltonian (\autoref{matrix}) in details. Divide it into three parts: time-dependent $(V_t)$, the part responsible for the interaction between the qubits $(V_J)$ and all the rest $(H_0)$.
\begin{equation}
	H^R=H_0+V_J+V_t
\end{equation}
The $\Omega_1\gg \Omega_2,\Delta_2\gg J$, so the driven field on the first qubit is strong and form the dressed states defined by the Hamiltonian $H_0$.
The Hamiltonian $H_0$ is station and we can find eigenvectors and eigenvalues. The drive on the second qubit is weak with very low intensity, so we can neglect any perturbation introduced by second drive on the position, width os the dressed states.
Firstly, we consider the system of non-interacting qubits $(J=0)$. In this case some transitions are prohibited and the splitting we are interested in is not observed. 
Therefore a non-zero interaction between qubits is necessary to observe AT effect.

After that according to the perturbation theory (with perturbation $V_J$) to $H_0$ we found the first-order corrections to the wave functions. We can use Fermi Golden Rule to find the probability of transition.
\appendix
\section{Theoretical description of two-photon process}
Perturbation theory allows semianalytical calculation of the transition rates (Rabi oscillation frequencies) for single and multi-photon processes.

Let's consider the Hamiltonian $H = H_0+V(t)$, with time-dependent interaction $V(t)$ and "reference" part -- $H_0$ and wave function $\Psi(t)$ . After switching to the interaction picture $V(t)$ modifies to $\tilde{V}(t) = \exp(i H_0 t)V(t)\exp(-i H_0 t)$ and Schrödinger equation gives by $i \derivative{\psi(t)}{t}=\tilde{V}(t)\psi(t)$, where $\psi(t)=\exp(i H_0 t)\Psi(t)$. The reference states of $H_0$ will be denoted $\ket{j}$, $H_0 \ket{j} = \omega_j \ket{j}$ whence it follows that matrix elements  of the modified interaction take the form $\bra{j}\tilde{V}(t)\ket{i}=\exp(i(\omega_j-\omega_i)t)\bra{j}V(t)\ket{i}$. A "solution" of Schrödinger equation in interaction picture corresponding to the initial reference state $\ket{i}$ is
\begin{equation}\label{forpsi}
	\psi_i(t)=\ket{i} - i \int_{-\infty}^{t}dt_1\tilde{V}(t_1)\ket{\psi_i(t_1)}
\end{equation}

The quantity of primary interest is not the wave function but the probability of transition from the initial state $\ket{i}\rightarrow\ket{f}$, where $\ket{f}$ is the final state of interest. Firstly, we find the transition amplitude $A_{fi} = \bra{f}U(t,t_0)\ket{i}=\bra{f}\psi_i(t)$ where $U(t,t_0)$ evolution operator.

Solving \autoref{forpsi} by simple iteration gives the series solution
\begin{equation}
	U(t,t_0) = 1+\sum_{n}U^{(n)}(t,t_0)
\end{equation}
\begin{equation}
	U^{(n)}(t,t_0) = -i^n \int_{t_0}^{t}d t_1\tilde{V}(t_1)...\int_{t_0}^{t_{n-1}}d t_n\tilde{V}(t_n)
\end{equation}
For "weak" interaction the series may be truncated to a fine term $n$ and the $n$th- order transition amplitude $\bra{f}U^{(n)}(t,t_0)\ket{i}$ can be evaluated. 
In our case the interaction in interaction picture satisfies the following relation
\begin{equation}
	\tilde{V}(t)=\exp(i H_0 t) \exp(-i H_0 t)
\end{equation}
